---
title: "mpatools"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tools for working with [Marine Protected Areas](https://www.iucn.org/theme/protected-areas/our-work/quality-and-effectiveness/world-database-protected-areas-wdpa) to compute species diversity metrics using [OBIS](https://obis.org/) data.


# Requirements

 + [R v4+](https://www.r-project.org/)
 
 + [rlang](https://CRAN.R-project.org/package=rlang)
 
 + [dplyr](https://CRAN.R-project.org/package=dplyr)
 
 + [readr](https://CRAN.R-project.org/package=readr)
 
 + [rappsdir](https://CRAN.R-project.org/package=rappsdir)

 + [wdpar](https://CRAN.R-project.org/package=wdpar)

 + [sf](https://CRAN.R-project.org/package=sf)
 
 + [vegan](https://CRAN.R-project.org/package=vegan)
 
 + [robis](https://cran.r-project.org/web/packages/robis/index.html)
 

# Installation

```
remotes::install_github("BigelowLab/mpatools")
```

# Usage


## Polygons for MPAs 

We use the [wdpar](https://CRAN.R-project.org/package=wdpar) R package to fetch MPAs by country (or 'global' for the complete dataset.)  By default, we store the files you download in the path returned by `rappdirs::user_data_dir("wdpar")` in [GeoPackage](https://www.geopackage.org/). Once you have downloaded format you can simply read the file from disk using `wdpa_read_country()`.

```{r fetch_country, message = FALSE}
library(rappdirs)
library(sf)
library(mpatools)
x <- wdpa_fetch_country(country = 'Cuba')
plot(sf::st_geometry(x))
```
