---
title: "mpatools"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tools for working with [Marine Protected Areas](https://www.iucn.org/theme/protected-areas/our-work/quality-and-effectiveness/world-database-protected-areas-wdpa) to compute species diversity metrics using [OBIS](https://obis.org/) data.


# Requirements

 + [R v4+](https://www.r-project.org/)
 
 + [rlang](https://CRAN.R-project.org/package=rlang)
 
 + [dplyr](https://CRAN.R-project.org/package=dplyr)
 
 + [readr](https://CRAN.R-project.org/package=readr)
 
 + [rappsdir](https://CRAN.R-project.org/package=rappsdir)

 + [wdpar](https://CRAN.R-project.org/package=wdpar)

 + [sf](https://CRAN.R-project.org/package=sf)
 
 + [vegan](https://CRAN.R-project.org/package=vegan)
 
 + [robis](https://cran.r-project.org/web/packages/robis/index.html)
 

# Installation

```
remotes::install_github("BigelowLab/mpatools")
```

# Usage


## Polygons for MPAs 

We use the [wdpar](https://CRAN.R-project.org/package=wdpar) R package to fetch MPAs by country (or 'global' for the complete dataset.)  By default, we store the files you download in the path returned by `rappdirs::user_data_dir("wdpar")` in [GeoPackage](https://www.geopackage.org/) format. Once you have downloaded a dataset you can simply read the file from disk using `read_mpa("name")`.  Below, we use the `list_mpa()` function to determine if we already have the Cuba dataset.

```{r fetch_country, message = FALSE}
library(rappdirs)
library(sf)
library(mpatools)
country <- "Cuba"
if (has_mpa(country)){
  mpa <- read_mpa(name = country)
} else {
  mpa <- fetch_mpa(name = country)
}
# 
plot(sf::st_geometry(mpa))
```

## OBIS occurrence data

Once you have MPA polygons in hand, it is easy to request OBIS records that fall within the polygons. 
We use the [robis](https://cran.r-project.org/web/packages/robis/index.html) R package to fetch occurrence data.  By default we store files you downlaod in the path returned by `rappdirs::user_data_dir("robis")` in [GeoPackage](https://www.geopackage.org/). Once you have downloaded format you can simply read the file from disk using `read_obis()`. 

```{r fetch_obis, message = FALSE, warning=FALSE}
if (has_obis(country)){
  obis <- read_obis(name = country, form = "sf")
} else {
  obis <- fetch_obis_country(name = country)
}
head(obis)
```

We actually find occurrences that fall within the convex hull of a set of polygons, but you can enforce a stricter policy using the `policy` argument for `fetch_obis()`.  See `?fetch_obis`.


```{r plot_obis}
plot(sf::st_geometry(mpa))
plot(sf::st_geometry(obis), col = 'orange', pch = 16, add = TRUE)
```

## Where are MPAs and OBIS data stored?

We use [rappsdir](https://CRAN.R-project.org/package=rappsdir) R package to manage the stirage location.  We store data here `rappdirs::user_data_dir("wdpar")` and here `rappdirs::user_data_dir("robis")`. This is a very effective way to store data **per user**.  But if you want to store data at a system-wide location you can do that using the ubiquitous `path` argument.

